\chapter{Prototype}
% \usemintedstyle{default}
\setlength{\DTbaselineskip}{14pt}
\begin{footnotesize}
  \dirtree{%
    .1 src/.
    .2 components/.
    .3 index.ts.
    .3 Viewer.tsx.
    .3 Navbar.tsx.
    .3 QueryPannel.tsx.
    .3 Queryinput.tsx.
    .3 Divider.tsx.
    .3 Button.tsx.
    .2 modules/.
    .3 atoms/.
    .4 index.ts.
    .4 cleanStart.ts.
    .4 endpoint.ts.
    .4 query.ts.
    .4 lru.ts.
    .4 ui.ts.
    .3 automations/.
    .4 index.ts.
    .4 useAutomations.ts.
    .4 BOT.ts.
    .4 GeoSPARQL.ts.
    .4 OBJ.ts.
    .3 viewer/.
    .4 index.ts.
    .4 types.ts.
    .4 useInitViewer.ts.
    .4 useLoadGeometry.ts.
    .3 fetchSPARQL.ts.
    .3 useCacheManagement.ts.
    .3 refTypes.ts.
    .2 pages/.
    .3 index.tsx.
  }
\end{footnotesize}

\setminted{
  linenos=true,
  bgcolor={white},
  style=default
}


\section{Components}
\subsection{Viewer}
\inputminted{tsx}{figures/snippets/src/components/Viewer.tsx}
\subsection{Navbar}
\inputminted{tsx}{figures/snippets/src/components/Navbar.tsx}
\subsection{QueryPannel}
\inputminted{tsx}{figures/snippets/src/components/QueryPannel.tsx}
\subsection{Queryinput}
\inputminted{tsx}{figures/snippets/src/components/Queryinput.tsx}
\subsection{Divider}
\inputminted{tsx}{figures/snippets/src/components/Divider.tsx}
\subsection{Button}
\inputminted{tsx}{figures/snippets/src/components/Button.tsx}

\section{Modules}
\subsection{Atoms}
\subsubsection{cleanStart}
\inputminted{ts}{figures/snippets/src/modules/atoms/cleanStart.ts}
\subsubsection{endpoint}
\inputminted{ts}{figures/snippets/src/modules/atoms/endpoint.ts}
\subsubsection{query}
\inputminted{ts}{figures/snippets/src/modules/atoms/query.ts}
\subsubsection{lru}
\inputminted{ts}{figures/snippets/src/modules/atoms/lru.ts}
\subsubsection{ui}
\inputminted{ts}{figures/snippets/src/modules/atoms/ui.ts}
\subsection{Automations}
\subsubsection{useAutomations}
\inputminted{ts}{figures/snippets/src/modules/automations/useAutomations.ts}
\subsubsection{BOT}
\inputminted{ts}{figures/snippets/src/modules/automations/BOT.ts}
\subsubsection{GeoSPARQL}
\inputminted{ts}{figures/snippets/src/modules/automations/GeoSPARQL.ts}
\subsubsection{OBJ}
\inputminted{ts}{figures/snippets/src/modules/automations/OBJ.ts}
\subsection{Viewer}
\subsubsection{types}
\inputminted{ts}{figures/snippets/src/modules/viewer/types.ts}
\subsubsection{useInitViewer}
\inputminted{ts}{figures/snippets/src/modules/viewer/useInitViewer.ts}
\subsubsection{useLoadGeometry}
\inputminted{ts}{figures/snippets/src/modules/viewer/useLoadGeometry.ts}
\subsection{fetchSPARQL}
\inputminted{ts}{figures/snippets/src/modules/fetchSPARQL.ts}
\subsection{useCacheManagement}
\inputminted{ts}{figures/snippets/src/modules/useCacheManagement.ts}
\subsection{refTypes}
\inputminted{ts}{figures/snippets/src/modules/refTypes.ts}

\section{Pages}
\subsection{index}
\inputminted{tsx}{figures/snippets/src/pages/index.tsx}
